---
title: "corrr"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

corrr is a package for making it easy to conduct and interpret correlations.

You can install:

- the latest development version from github with

```{r install_git, eval = FALSE}
if (packageVersion("devtools") < 1.6) {
  install.packages("devtools")
}
devtools::install_github("drsimonj/corrr")
```

## Getting Started

The central aspect of corrr is working with a data frame of correlations. The primary function of corrr is `cor_frame()`, which extends `stats::cor()` by returning the correlations (`r`) and corresponding sample sizes (`n`) of each variable pair (`x` and `y`) in a `dplyr::data_frame()`.

```{r}
library(MASS)
library(corrr)

# Simulate four columns correlating about .7 with each other
mu <- rep(0, 4)
Sigma <- matrix(.7, nrow = 4, ncol = 4) + diag(4)*.3
set.seed(1) 
d <- mvrnorm(n = 1000, mu = mu, Sigma = Sigma)

# Insert a few missing values
d[100:600, 1] <- NA
d[320:322, 2] <- NA

# Use corrr::cor_frame() to obtain a data frame of the correlations
cor_frame(d)
```

We also have the option to gain the same information in a list of two matrix objects using `cor_matrix()`

```{r}
cor_matrix(d)
```
