---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

---
title: "corrr"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

corrr is a package for exploring **corr**elations in **R**. It makes it possible to easily perform routine tasks when exploring correlation matrices such as ignoring the diagonal, focusing on the correlations of certain variables against others, or arranging and visualising the matrix in terms of the strength of the correlations.

You can install:

- the latest development version from github with

```{r install_git, eval = FALSE}
if (packageVersion("devtools") < 1.6) {
  install.packages("devtools")
}
devtools::install_github("drsimonj/corrr")
```

## Using corrr

Using `corrr` starts with `correlate()`, which acts like the base correlation function `cor()`. It differs by defaulting to pairwise deletion, and returning a correlation data frame (`cor_df`) of the following structure:

- A `tbl` with an additional class, `cor_df`
- An extra "rowname" column
- Standardised variances (the matrix diagonal) set to missing values (`NA`) so they can be ignored.

### API

The corrr API is designed with data pipelines in mind (e.g., to use `%>%` from the magrittr package). After `correlate()`, the primary corrr functions take a `cor_df` as their first argument, and return a `cor_df` or `tbl` (or output like a plot). These functions serve one of three purposes:

Internal changes (`cor_df` out):

- `shave()` the upper or lower triangle (set to `r NA`).
- `rearrange()` the columns and rows based on correlation strengths.

Reshape structure (`tbl` or `cor_df` out):

- `focus()` on select columns and rows.
- `stretch()` into a long format.

Output/visualisations (console/plot out):

- `rplot()` plots the correlations.

## Examples

```{r example, message = FALSE, warning = FALSE}
library(MASS)
library(corrr)
set.seed(1)

# Simulate three columns correlating about .7 with each other
mu <- rep(0, 3)
Sigma <- matrix(.7, nrow = 3, ncol = 3) + diag(3)*.3
seven <- mvrnorm(n = 1000, mu = mu, Sigma = Sigma)

# Simulate three columns correlating about .4 with each other
mu <- rep(0, 3)
Sigma <- matrix(.4, nrow = 3, ncol = 3) + diag(3)*.6
four <- mvrnorm(n = 1000, mu = mu, Sigma = Sigma)

# Bind together
d <- cbind(seven, four)
colnames(d) <- paste0("v", 1:ncol(d))

# Insert some missing values
d[sample(1:nrow(d), 100, replace = TRUE), 1] <- NA
d[sample(1:nrow(d), 200, replace = TRUE), 5] <- NA

# Correlate
x <- correlate(d)
class(x)
x
```

As a `tbl`, we can use functions from data frame packages like `dplyr`, `tidyr`, `ggplot2`:

```{r, message = FALSE, warning = FALSE}
library(dplyr)

# Filter rows by correlation size
x %>% filter(v1 > .6)
```

corrr functions work in pipelines (`cor_df` in; `cor_df` or `tbl` out):

```{r combination, warning = FALSE, fig.height = 4, fig.width = 5}
datasets::mtcars %>%
  correlate() %>%    # Create correlation data frame (cor_df)
  focus(-cyl, -vs, mirror = TRUE) %>%  # Focus on cor_df without 'cyl' and 'vs'
  rearrange(method = "HC", absolute = FALSE) %>%  # arrange by correlations
  shave() %>%  # Shave off the upper triangle for a clean plot
  rplot()  # Plot the results
```